---
title: "Case Financials at Carco"
author: "Caio Stabel & Rodrigo Giannotti"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magrittr)
library(readxl)
```

# Introdução

You are the CFO for Carco, a small car rental company. 
You are trying to get some idea of what Carco’s financial and income statements will look like during the fixed year (year 0) and the next five years.

As variáveis e suas relações são listadas na primeira página do documento e serão criadas por tópico a seguir.

## Variables that are fixed

```{r}
tax_rate <- 0.470
dividend_rate <- 0.700
retention_rate <- 0.300
intrest_rate_old <- 0.105
interest_rate_new <- 0.095
repayment_rate <- 0.200
```


## Factors

## Sales

```{r}
sales_factor <- c(0,0,0,0,0,0)
sales_mean <- 1.10
sales_sd <- 0.05

sales <- c(1000,0,0,0,0,0)

for (year in 2:6)) {
  sales[year] <- sales[year] * sales[year]
}
```

### Current Liabilities

```{r}
current_liabilities_factor <- c(0,0,0,0,0,0)
current_liabilities_mean <- 0.08
current_liabilities_sd <- 0.01

current_liabilities <- c(0,0,0,0,0,0)

for (year in 1:6)) {
  current_liabilities[year] <- current_liabilities_factor[year] * sales[year]
}
```

### Current Assets

```{r}
current_assets_factor <- c(0,0,0,0,0,0)
current_assets_mean <- 0.15
current_assets_sd <- 0.02

current_assets <- c(0,0,0,0,0,0)

for (year in 1:6)) {
  current_assets[year] <- current_assets_factor[year] * sales[year]
}
```

### Net Fixed Assets

```{r}
fixed_assets_factor
fixed_assets_mean <- 0.77
fixed_assets_sd <- 0.04

fixed_assets <- c(0,0,0,0,0,0)

for (year in 1:6)) {
  fixed_assets[year] <- fixed_assets_factor[year] * sales[year]
}
```

### Expense

```{r}
expense_factor
expense_mean <- 0.80
expense_sd <- 0.06

expense <- c(0,0,0,0,0,0)

for (year in 1:6)) {
  expense[year] <- expense_factor[year] * sales[year]
}
```

### Total Assets

```{r}
total_assets = fixed_assets + fixed_assets
```

### Depreciation

```{r}
depreciation <- c(0,0,0,0,0,0)

acc_depreciation <- c(0,0,0,0,0,0)
acc_depreciation[1] <- 330

for (year in 2:6)) {
  depreciation[year] <- acc_depreciation[year - 1] + (0.1 * fixed_assets_at_cost[year - 1])
  acc_depreciation[year] <- acc_depreciation[year - 1] + depreciation[year]
}
```

### Fixed Assets at Cost

```{r}
for (year in 1:6)) {
  fixed_assets_at_cost[year] <- depreciation[year] + fixed_assets[year]
}
```









